### 2013-05-16 paper/paper.tex
Minproved section.

% TEMPLATE for Usenix papers, specifically to meet requirements of
%  USENIX '05
% originally a template for producing IEEE-format articles using LaTeX.
%   written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his excellent SWIG paper in Proceedings,
%   Tcl 96
% turned into a smartass generic template by De Clarke, with thanks to
%   both the above pioneers
% use at your own risk.  Complaints to /dev/null.
% make it two column with no page numbering, default is 10 point

% Munged by Fred Douglis <douglis@research.att.com> 10/97 to separate
% the .sty file from the LaTeX source template, so that people can
% more easily include the .sty file into an existing document.  Also
% changed to more closely follow the style guidelines as represented
% by the Word sample file.

% Note that since 2010, USENIX does not require endnotes. If you want
% foot of page notes, don't include the endnotes package in the 
% usepackage command, below.

% This version uses the latex2e styles, not the very ancient 2.09 stuff.
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{epsfig,graphicx,usenix,fullpage, hyperref}

%\usepackage{endnotes}
\begin{document}

%don't want date printed

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf 6.829 Final Project: A comparison of bit-rate selection algorithms}
%for single author (just remove % characters)
\author{
{\rm Colleen Josephson}\\
cjoseph@mit.edu
\and
{\rm Pavel Panchekha}\\
pavpan@mit.edu
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

\date{May 2013}

\maketitle

% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}

\subsection*{Abstract}
In this paper we compare the performance SampleRate and Minstrel, two popular bit-rate selection algorithms that have widespread real-world usage on the Linux operating system. We use a novel trace-based approach that allows us avoid in-depth kernel programming and write simple Python scripts instead. We also introduce improvements to the Minstrel algorithm that allow for significant gains in throughput. 
\section{Introduction}

One of the key ways that wireless networks differ from wired is that wireless networks have varying link rates. Link conditions vary with time due to interference from other devices, changing network geometry, and mobile clients. An optimal rate at some time $t$ may be different from the optimal rate just 30 seconds earlier. A good bit-rate selection algorithm has to detect and adapt to these conditions. If the chosen bit-rate is too slow, then the throughput will be unnecessarily low. If the rate is too high, then failures will be very frequent and throughput will again suffer.

There are three main classes of bit-rate adaptation protocols: frame-based, SNR-based, and cross-layer protocols. Frame-based protocols measure the fraction of successfully received packets. SNR protocols make decisions based on the estimated Signal to Noise ratio. Cross-layer protocols use SoftPHY data from the physical layer. The most commonly implemented protocols on today's networks are frame based, because SNR protocols perform poorly~\cite{samplerate}, and cross-layer protocols cannot be deployed on current networks as they violate the network layering abstraction.

The two most popular frame-based protocols are SampleRate and Minstrel. Both were implemented in the MadWifi drivers for Linux wireless, and today Minstrel is the default bit-rate selection algorithm for all wireless drivers on Linux. Our research used a trace-based approach to analyze the performance of these two algorithms. We also created a modified version of Minstrel that provides significant throughput gains over the vanilla Minstrel implementation.

This paper will provide a general overview of SampleRate and Minstrel. We then discuss our testing methodology. We compare the performance of the two algorithms, and then discuss modifications we made to Minstrel and compare it's performance to the current kernel implementation.
\section{SampleRate}

SampleRate has three main functions: \texttt{ApplyRate()}, which returns a the bit-rate to send a packet at, \texttt{ProcessFeedback()}, which updates the statistics for a bitrate after an attempted packet send,  and \texttt{RemoveStaleResults()}, which removes all statistics older than 10 seconds.

SampleRate sends at the rate with the lowest average transmission time. Every ten packets, it sends out a sample packet at a random rate that has a smaller predicted transmission time than the current rate. SampleRate updates its transmission time estimates every ten seconds based on feedback from the card about which packets succeeded and at which rate they were sent. If the current best rate experiences four successive failures, SampleRate stops using that rate.

\section{Minstrel}

Minstrel utilizes the multi-rate retry chain (mrr), which is an array that tells the card which rates to try before reporting a failure. The retry chain makes failures extremely unlikely. Minstrel has a mrr of [\emph{highest throughput, next highest throughput, best probability, lowest base rate}] for most packets. In essence, Minstrel tells the card to send at the highest throughput rates but to fall back to reliable rates upon failure. The throughput and success probability are calculated for each rate every 100ms. 
To keep an accurate estimate of throughput and probability for each rate, Minstrel sends sample frames ten percent of the time. If the randomly chosen rate has a higher lossless throughput than the current optimal rate, the mrr is as follows: [\emph{random, highest throughput, best probability, lowest base rate}]. If the random rate is slower than the optimal rate, it is moved further back in the retry chain: [\emph{best throughput, random, best probability, lowest base rate}]. This ensures that Minstrel never samples rates worse than the current optimal rate unless the optimal rate experiences a failure. See table ~\ref{table:1} for an overview of the mrr's used in Minstrel.

Minstrel's implementation is based on SampleRate's, so it also has \texttt{ApplyRate()} and \texttt{ProcessFeedback()}, as well as a \texttt{UpdateStats()} method that runs every 100ms. \texttt{UpdateStats()} is home to some of the key differences between Minstrel and SampleRate. Instead of making decisions based on the average 
transmission time, Minstrel uses throughput: \begin{equation} T = \frac{p}{tx\_time}\end{equation} where $T$ is the throughput of some rate $r$, $p$ is the probability of successfully transmitting at $r$, and $tx\_time$ is the lossless transmission time at $r$.

The probability $p$ of a successful transmission is calculated from statistics collected by the \texttt{ProcessFeedback()} method, and is done using an exponential weighted moving average, or EWMA. The EWMA creates a weighted average that weighs recent data more heavily than old data. The weighting for old data decreases exponentially. An EWMA is beneficial because it ensures that a sudden degradation in link quality will create a rapid response in the probabilities due to the increased weight of recent data. Minstrel will waste less time sending packets at a rate that no longer works, as compared to a simple probability calculation.

\begin{table*}[htb]
  \centering
    \begin{tabular}[htb]{l|l|l|l}
    \textbf{Try} & \textbf{Normal}               & \textbf{Random (slower than best)} & \textbf{Random (faster than best)} \\ \hline
    1   & Best throughput      & Best throughput           & Random                    \\
    2   & Next best throughput & Random                    & Best throughput           \\
    3   & Highest probability  & Highest probability       & Highest probability       \\
    4   & Lowest rate          & Lowest rate               & Lowest rate               \\
    \end{tabular}
    \caption{Multi-rate retry chains for Minstrel}
\label{table:1}
\end{table*}

\section{Methodology}

We collected traces about the success rates of each bit rate through a modified Linux wireless driver; Minstrel and SampleRate were then re-implemented in Python and replayed on this collected data.

\subsection{Motivation}

Our initial plan was to use the algorithms as implemented in the MadWifi drivers for the Atheros chipset. However, MadWifi is deprecated, and only runs on now-ancient network cards. In addition to the unavailability of compatible hardware, MadWifi would not be an accurate reflection of contemporary users. Both ath5k and ath9k, which were created to replace MadWifi, did not port the old bit-rate selection system from MadWifi and instead opted to use the Linux kernel's rate selection framework, which only implements Minstrel.  Porting SampleRate over to the new drivers would not have been trivial, as there are vast differences in the assumed interfaces.

At this point, we decided to take a different approach and use an approach similar to that used in Sprout~\cite{sprout}.  We modified the ath9k driver to try bit rates uniformly and report packet successes and failures, then analyze this data in user-space with re-implementations of SampleRate and Minstrel.  Not only did this lessen the re-implementation cost (since the algorithms were re-implemented in user-space in Python, not in kernel-mode in C), but it also provided reproducibility and allowed us to investigate modifications to SampleRate and to Minstrel.

\subsection{Trace Collection}

Traces were collected from an ath9k driver modified to sample bit-rates uniformly (forgoing the usual bit rate selection algorithm) and to report successes and failure from each packet.  Each bit-rate was sampled uniformly in time -- thus, more packets were sent at higher bit-rates, since those bit-rates otherwise deliver packets faster.  For each packet sent, the time to send was recorded, and the bit rate for the next packet was chosen to be the bit-rate that had the smallest total transmissions time.  This guarantees approximately uniform selection of bit-rates.  The wireless card used was an Qualcomm Atheros \texttt{AR9285} Wireless Network Adapter.

Traces were collected at an unused access point in a basement of MIT.  Before initiating trace collection, the wireless card was put in monitor mode for a minute to check that no other users were connected to the access point.  None were, guaranteeing that traces would not include delays from carrier sensing or run into problems due to congestion at the access point.  Traffic was simulated by sending 1500 byte packets to the access point's IP address.  Packets were sent over UDP to avoid the TCP congestion window and retry loop; since no users were using the access point, there was no real risk of deteriorating service for other users.  Figure~\ref{figure:1} shows a floor plan of the access point and basement.

Eighteen total traces were collected, grouped into a number of scenarios.  Each trace was 30 seconds to a minute long, and each scenario was repeated twice.  One scenario placed the wireless card with line of sight to the access point, two meters from it.  Another created solid body interference by having one of the authors stand between the wireless device and the access point.  One placed the device around a corner from the access point, creating some potential for multi-path interference and fading.  In one scenario, we moved the wireless device throughout the trace, pacing back and forth in a 10 meter line in front of the access point.  Finally, one five-minute trace was taken, which mixed the above scenarios, first moving around a corner, then standing still, then moving to line-of-sight with the access point.  A few shorter (3 to 15 second long) traces were collected to investigate the startup behavior of bit-rate selection.

\begin{figure}[hbt]
  \centering
  \includegraphics[width=3.5in]{bsmt13floorplan.png}
  \caption{We collected most traces at a remote access point in the basement of Building 13 at MIT.}
\label{figure:1}
\end{figure}

\subsection{Testing framework}

Once collected, traces were fed to a bit rate selection simulator to analyze the performance of Minstrel and SampleRate.  Both algorithms were represented by two functions: \texttt{ApplyRate}, which returned a multi-rate retry chain of rates to attempt transmission at, and \texttt{ProcessFeedback}, which was passed the success or failure of the transmission and how many attempts were made at each bit rate (along with the current time and the total time to send the packet); this parallels the interface used by the Linux kernel itself.  Helper functions, like SampleRate's \texttt{RemoveStaleResults} function, were instead called from one of \verb|apply_rate| or \verb|process_feedback|.  For every packet transmission, the simulator would call \texttt{ApplyRate}; compute the probability of success based on packets sent in 10-millisecond window around the current time; and from that determine the success or failure of the packet, and the number of retries required.  The results of a test run are thus non-deterministic; both algorithms tested are non-deterministic anyway, so this is no handicap.  To improve re-implementation fidelity, both implementations were meticulously checked by both authors.

We implemented SampleRate as outlined in John Bicket's Master's thesis~\cite{samplerate}. This implementation is slightly different from how it was implemented in the MadWifi kernel driver. The primary difference is that the kernel implementatiion uses a EWMA, while the thesis implementation computes average transmission times based on a 10-second sliding window. It would be worth looking at the performance of EWMA-based SampleRate in the future.  Our implementation, \texttt{samplerate.py} contains \texttt{ApplyRate()}, \texttt{ProcessFeedback()} and  \texttt{RemoveStaleResults()}, as well as a few helper functions and data structures for tracking rate statistics.

We implemented Minstrel by porting the C code from the 3.3.8 version of the Linux kernel into Python. \texttt{minstrel.py} contains \texttt{ApplyRate()}, \texttt{ProcessFeedback()} and  \texttt{UpdateStats()}, as well as a few helper functions and data structures for tracking rate statistics.

\subsection{Experimental parameters}

There are a few parameters that can be adjusted in Minstrel and SampleRate, but our implementation is consistent with what the respective authors originally chose.

In SampleRate, the number of successive retries, the frequency of probe packets, the window size, and the lossless $tx\_time$ estimations are all parameters that can be changed. The SampleRate thesis chose 4 successive retries, a probe frequency of every ten packets, and a window size of ten seconds. We used the same values in our implementation. Additionally, we use the same equation to calculate $tx\_time$.

In Minstrel, the frequency of probe packets, the window size, and the EWMA weighting are the adjustable parameters. The 3.3.8 Linux kernel uses a probe frequency of 10\%, a window size of 100ms, and a EWMA weighting of 0.75. We use the same values. Additionally, Minstrel also estimates $tx\_time$ using the \texttt{ieee80211\_frame\_duration()} method from the \texttt{util.c} file in \texttt{net/mac80211}. We implemented the same function in Python.

\section{Analysis} \label{sec:analysis}

Minstrel and SampleRate perform similarly, usually within 3 Mbps of each other. In especially noisy situations, such as when the client is moving or when the client is around the corner from the access point, Minstrel consistently outperforms SampleRate. Figure~\ref{figure:3} shows the results of a single run of both Minstrel and SampleRate over all the different types of traces we collected. This is consisten with the literature---the authors of Minstrel noted that they focused on making Minstrel robust to poor conditions~\cite{minstrel}, which means losing potential througput gains in stable situations. Conversely, John Bicket noted that SampleRate performs better in stable situations than noisy or highly varying conditions~\cite{samplerate}. We were slightly surprised to see how significantly SampleRate outperformed Minstrel in certain situations, though, with throughput being up to $67\%$ higher.

In Figure~\ref{figure:2} we show the performance of constant bitrates on the \texttt{clear\_1.dat} trace, which is a 35 second trace in clear view of the access point. The optimal rate here was 18 Mbps, which achieved an average throughput of about 15 Mbps. The next highest bitrate, 24 Mbps, shows a large plummet in performance, down to a throughput of 1.17 Mbps. This steep drop is charateristic, and is also talked about in the SampleRate thesis~\cite{samplerate}. It demonstrates just how important bit rate selection is---the optimal bit rate has typically much higher throughput than the next-best rate, but if the algorithm strays too high the results can be devastating.

Bit rate selection algorithms strive to have a throughput near or greater than the optimal constant rate. For this \texttt{clear\_1.dat} trace, SampleRate clearly meets this goal, achieving an average throughput more than 1 Mbps higher than the optimal constant rate. In noisier situations, though, Minstrel tends to be closer to the optimal constant rate.

We did notice that Minstrel occasionally has very poor performance on traces 1 seconds or shorter---sometimes as low as 2 Mbps. This is because it does not stop sending after four successive failures, but must wait until the end of the 100ms window for statistics to be compiled. Thus it wastes a lot of time sending at non-optimal rates; and a short trace has only a few 100ms windows. This behavior is not evident in long traces, however, as Minstrel has had time to both smooth out bad windows and collect more statistics (making it less likely to choose a bad rate).

Bit rate selection algorithms naturally show variation between runs on the same data because probe rates are chosen randomly. The variations were small, though. On ten runs on the \texttt{clear\_1.dat} trace, the average of Minstrel over 10 runs was 7.52 Mbps, with a maximum variation of +/- 0.12 Mbps. SampleRate was similarly stable, it averaged 5.284 Mbps over 10 runs, with a maximum variation of about +/- 0.3 Mbps. The variation of results on other traces (except the very short traces, where variance was naturally higher) was similar.

The histograms of which rates the algorithms decided to send at provided some useful insight into their behavior. Table~\ref{table:2} is a histogram of a 35 second corner trace. In this situation Minstrel outperforms SampleRate. The constant rate throughputs demonstrate that 11, 12 and 18 Mbps were the optimal rates to send at, and that no packets were sucessful at any higher rates. SampleRate does not use these rates except for probe packets, which happens once every ten seconds. Minstrel, since it does not terminate transmission after four successive failures, spends a lot of time sending probe packets at these high rates, which are never successful. The only reason low rates have fewer probe packets is because Minstrel explicitly places them second in the retry chain, and because the main rate usually succeeds. Despite the fact that SampleRate spends less time probing high rates, it is less successful. This is because Minstrel send the most packets at the three most successful bitrates. SampleRate, however, spends too much time sampling low bit rates. Additionally, it spends a lot of time sending at 5.5 Mbps. TODO: why does SampleRate spend time at 5.5 Mbps??? We suspect that this might be because the equation SampleRate uses to estimate average transmission time needs to be tweaked.

Our examination of the histograms showed us that Minstrel spends a lot of time sampling at high rates. In the specific case we examined, they were all probe packets. But sometimes Minstrel would choose 54Mbps as the rate with the best throughput, which made little sense. A closer investigation revealed that Minstrel would sometimes get lucky with a successful probe at 54Mbps. The probability of success would be 1, so then 54Mbps would be chosen as the rate with the best throughput. This is unfortunate, but the real harm comes in the next window: packets would now be send at 54Mbps, and nearly all would fail.  However, since these hundreds of failed packets weight equally with the single successful probe packet, Minstrel still considers 54 Mbps the best rate. Thus the probability of success might drop from 1 to about 0.70, which is still high enough for the throughput equation to rank 54 Mbps highly. Fundamentally, the problem is that the EWMA does not account the second block having hundreds more packets than the first.  In the next section we will talk about Minproved, an improved version of Minstrel where we attempt to fix this problem.

\begin{figure}[htb]
  \hspace{-1em}\includegraphics[width=3.4in]{constant.png}\vspace{-0em}
  \caption{Throughput of using a constant 802.11b/g bitrate in clear line-of-sight from the access point.}
\label{figure:2}
\end{figure}

\begin{figure}[htb]
  \hspace{-.75em}\includegraphics[width=3.5in]{MinVSSam1.png}\vspace{-0.75em}
  \caption{Throughput of Minstrel and SampleRate on various trace situations.}
\label{figure:3}
\end{figure}

\section{Improvements to Minstrel}

In Section~\ref{sec:analysis}, we noted that the EWMA used by Minstrel causes problems when probe packets succeed.  A natural next step was to modify Minstrel to avoid this problem, by weighing each 100ms block of statistics in proportion to how many packets it contains.  To preserve the exponential weighing of statistics, we chose to weigh each 100ms block based on the number of packets sent, compared to the those sent in the average block.  Parameters were chosen so that if all blocks were of average size, our ``balanced'' EWMA would behave identically to a normal EWMA.

To achieve, this weighing, we first recast the usual EWMA algorithm.  The usual algorithm computes $$p \gets \alpha p + (1 - \alpha) \frac{n_i}{d_i}),$$ where $p$ is the EWMA probability, $\alpha$ is a weight factor equal to $0.75$ in Minstrel, and $n_i$ and $d_i$ are the number of successful and total packets in the 100ms window.  We next write this as $$p \gets \frac{\beta p + \frac{n_i}{d_i}}{\beta + 1},$$ where $\beta = \alpha / (1 - \alpha)$.  We can now change the factor in front of $n_i / d_i$ to account for the total number of packets sent.  In our case, we chose to compute $$p \gets \frac{\beta p + w_i \frac{n_i}{d_i}}{\beta + w_i},$$ where the weight $w_i$ is $d_i / (d / b)$, with $d$ the total number of packets ever send and $b$ the total number of 100ms windows ever seen.  Then $d / b$ is the average number of packets per window and $d_i / (d / b)$ is the ratio between the current window and the average window.  Note that if the current window is of average size, this is exactly equal to the usual EWMA result.  Finally, we can simplify the above equation to $$S p \gets \frac{\beta \frac{d}{b} p + S n_i}{\beta \frac{d}{b} + d_i},$$ where $S$ is any constant.  This equation is usable in fixed-precision arithmetic, making it simple to implement in the Linux kernel (which prefers to avoid floating point computations for portability reasons).  Because packets are high rates rarely succeed, the average window size is a few packets; thus, the main qualitative difference between our balanced EWMA and the original Minstrel EWMA is that non-probe packets at probe bit-rates are weighed extraordinarily heavily, and this fact allows us to avoid multiple failing 100ms windows.

We tested a modified Minstrel which used this more balanced EWMA algorithm on the same traces as Minstrel and SampleRate, and found a consistent improvement of approximately 1 Mbps over the normal Minstrel algorithm.

We noted another curious fact about Minstrel: the kernel implementation of Minstrel samples rates very frequently.  This is due to a flag the kernel uses to track which probe bit-rates were actually sampled.  In cases where the probe packets are at a lower bit-rate than the current bit-rate, the probe packets might never be sent if the current bit-rate succeeds.  However, the flagging seems to misbehave in certain cases, causing the extremely aggressive probing we see in some traces.  For example, in one of the traces where the client was around a corner from the access point, Minstrel sent almost half of its packets at probe frequencies.  We modified Minstrel to be less aggressive with probe packets, decreasing the sampling frequency to approximately 10\%.  This lead to more improvement, especially on noisy traces such as the cases where the client was around a corner or moving.

Overall, our improved Minstrel, which we call Minproved, often out-performs Minstrel by multiple Mbps and never doing any worse; and usually surpassing the best constant bit rate.  In only one case did SampleRate achieve higher throughput than Minproved, and in this case both surpassed the best constant bit-rate.  In cases where Minstrel out-performed SampleRate, the same behavior was true of Minproved.  In most cases, Minproved was $30\%$ faster than Minstrel, and in some the improvement as great as $60\%$.

However, Minproved still makes many poor choices.  While a single successful probe packet does not cause many hundreds of milliseconds of failed packets, as it does in Minstrel, the single successful packet can still cause one window's worth of failed packets.  Instead, it would be best if a successful probe packet at a rarely-successful bit-rate was treated more carefully.  One can imagine an algorithm, for example, which would treat a single successful probe packet as a reason to send a dozen

\begin{table*}[htb]
    \centering
    \begin{tabular}{l|l|l|l|l}
    \textbf{Rate in Mbps}     & \textbf{SampleRate} & \textbf{Minstrel}   & \textbf{Minproved}   & \textbf{Constant Rate Throughput}\\ \hline
    1                & 631        & 39         & 29      & 0.639 Mbps    \\
    2                & 863        & 82         & 20      & 1.425 Mbps    \\
    5.5              & 4814       & 110        & 805     & 4.406 Mbps    \\
    6                & 423       & 137        & 276     & 4.603 Mbps    \\
    9                & 553         & 327        & 57      & 4.630 Mbps   \\
    11               & 3326       & 4121       & 3356    & 9.627 Mbps   \\
    12               & 4649       & 9515       & 14285   & 9.444 Mbps   \\
    18               & 1154         & 13295      & 15795   & 8.458 Mbps    \\
    24               & 16         & 6909       & 1918    & 0 Mbps   \\
    36               & 16         & 6944       & 2030    & 0 Mbps   \\
    48               & 16         & 6888       & 2009    & 0 Mbps   \\
    54               & 16         & 7308       & 1953    & 0 Mbps   \\ \hline
    \textbf{Avg. Throughput:} & \emph{5.42 Mbps}  & \emph{7.578 Mbps} & \emph{11.107 Mbps} \\
    \end{tabular}
    \caption{Histograms of the number of packets sent at each rate.  The trace was \texttt{corner\_1.dat} and was recorded around the corner from the access point. The best fixed rate for this trace was 11 Mbps, which achieved a throughput of 9.627 Mbps. Minproved achieves a higher throughput than the best fixed rate.}
\label{table:2}
\end{table*}

\begin{figure}[htb]
  \hspace{-2em}\includegraphics[width=3.5in]{mnVSspVSmp1.png}\vspace{-0em}
  \caption{Throughput of Minproved as compared to SampleRate and vanilla Minstrel.}
\label{figure:4}
\end{figure}

\subsection{Availability}
All of our code and collected trace data are available on GitHub:

\noindent
{\small\url{https://github.com/pavpanchekha/6.829-project/tree/3.8.6}}

This repository contains the modified ath9k driver used to collect the traces, the traces we collected, as well as the Python simulation framework.

\section{Future Work}

Very little analysis has been done of the performance of bit-rate selection algorithms on 802.11n networks. 802.11n introduces many new rates, as well as multiple-input/multiple output, or MIMO, capabilities. There is an 802.11n version of Minstrel, but there is no implementation of SampleRate.

We implemented SampleRate as outlined in the thesis, but the MadWifi implementation has a few key differences from the thesis. We suspect that the MadWifi use of an EWMA instead of a window has more accurate rate statistics, which will potentially lead to better throughput. We did not have the time to implement this alternative version, but it is worth further examination.

Hari Balakrishnan suggested that we use a broadcast approach when collecting our traces, instead of link layer acknowledgments. This would entail having multiple computers listening at the broadcast address, and compiling successes and failures from that information. This approach would free us from certain idiosyncrasies of reading data from the driver, and lead to possibly more accurate traces. We were unable to implement this due to a lack of time and ability to acquire compatible hardware.

Finally, we are considering submitting our improvements to the Minstrel algorithm, namely the BEWMA, as a kernel patch. Currently is implemented in Python, so we would have to port the code into C.

\section{Conclusion}

Minstrel and SampleRate perform similarly. The results suggest that SampleRate tends to do slightly better, except in very lossy links such as the client being far awat from the AP or actively moving. Minproved, our improved version of Minstrel, solidly outperforms both Minstrel and SampleRate.

It seems like the abandonment of SampleRate in the Linux kernel may have been a bit hasty, since it seems to perform similarly to Minstrel and it is a much simpler algorithm and thus less bug prone. We are not sure if there are adjustments to SampleRate that will bring it's performance near Minproved, though.

\section{Acknowledgments}
We would like to thank Jonathan Perry and Hari Balakrishnan for their valuable guidance, as well as Derek Smithies for answering our questions about Minstrel.

{\footnotesize \bibliographystyle{acm}
\bibliography{paper}}

\end{document}