### 2013-05-16 paper/paper.tex
TeX cleanup

% TEMPLATE for Usenix papers originally for producing IEEE-format articles
% written by Matthew Ward, CS Department, Worcester Polytechnic Institute.
% adapted by David Beazley for his paper in Proceedings, Tcl 96
% turned into a smartass generic template by De Clarke

\documentclass[twocolumn,10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx,hyperref}
\usepackage{subcaption}

\title{6.829 Final Project: A comparison of bit-rate selection algorithms}
\author{Colleen Josephson \\ \texttt{cjoseph@mit.edu}
  \and Pavel Panchekha \\ \texttt{pavpan@mit.edu}}
\date{May 2013}

\begin{document}
\maketitle

\begin{abstract}
We compare the performance SampleRate and Minstrel, two popular
bit-rate selection algorithms that have widespread real-world usage.
We use a trace-based approach to avoid kernel programming and consider
Python implementations of both SampleRate and Minstrel.  We test both
algorithms in multiple real-world scenarios, including scenarios with
mobile clients and noisy environments, to highlight differences
between the two algorithms.  We also introduce improvements to the
Minstrel algorithm that allow for significant gains in throughput.
\end{abstract}

\section{Introduction}

One of the key ways that wireless networks differ from wired links is
that wireless networks have varying link rates.  Link conditions vary
with time due to interference from other devices, changing network
geometry, and mobile clients.  An optimal rate at one time may be
different from the optimal rate just 30 seconds earlier.  A good
bit-rate selection algorithm has to detect and adapt to these
conditions.  If the chosen bit-rate is too slow, then the throughput
will be unnecessarily low; if the rate is too high, failures will
be very frequent and throughput will again suffer.

There are three main classes of bit-rate adaptation protocols:
frame-based, SNR-based, and cross-layer protocols.  Frame-based
protocols measure the fraction of successfully received packets.  SNR
protocols make decisions based on the estimated Signal to Noise
ratio.  Cross-layer protocols use SoftPHY data from the physical
layer.  The most commonly implemented protocols on today's networks are
frame based, because SNR protocols perform poorly~\cite{samplerate},
and cross-layer protocols cannot be deployed on current networks as
they violate the network layering abstraction.

The two most popular frame-based protocols are SampleRate and
Minstrel.  Both were implemented in the MadWifi drivers for Linux
wireless, and today Minstrel is the default bit-rate selection
algorithm for all wireless drivers on Linux.  Our research used a
trace-based approach to analyze the performance of these two
algorithms.  We also created a modified version of Minstrel that
provides significant throughput gains over the vanilla Minstrel
implementation.

This paper will provide a general overview of SampleRate and
Minstrel.  We then discuss our testing methodology.  We compare the
performance of the two algorithms, and then discuss modifications we
made to Minstrel and compare it's performance to the current kernel
implementation.

\section{SampleRate}

SampleRate has three main functions: \texttt{ApplyRate()}, which
returns a the bit-rate to send a packet at,
\texttt{ProcessFeedback()}, which updates the statistics for a bit-rate
after an attempted packet send, and \texttt{RemoveStaleResults()},
which removes all statistics older than 10 seconds.

SampleRate sends at the rate with the lowest average transmission
time.  Every ten packets, it sends out a sample packet at a random rate
that has a smaller predicted transmission time than the current
rate.  SampleRate updates its transmission time estimates every ten
seconds based on feedback from the card about which packets succeeded
and at which rate they were sent.  If the current best rate experiences
four successive failures, SampleRate stops using that rate.

\section{Minstrel}

Minstrel utilizes the multi-rate retry chain (mrr), which is an array
that tells the card which rates to try before reporting a failure.  The
retry chain makes failures extremely unlikely.  Minstrel has a mrr of
[\emph{highest throughput, next highest throughput, best probability,
    lowest base rate}] for most packets.  In essence, Minstrel tells
the card to send at the highest throughput rates but to fall back to
reliable rates upon failure.  The throughput and success probability
are calculated for each rate every 100ms.  To keep an accurate
estimate of throughput and probability for each rate, Minstrel sends
sample frames ten percent of the time.  If the randomly chosen rate has
a higher lossless throughput than the current optimal rate, the mrr is
as follows: [\emph{random, highest throughput, best probability,
    lowest base rate}].  If the random rate is slower than the optimal
rate, it is moved further back in the retry chain: [\emph{best
    throughput, random, best probability, lowest base rate}].  This
ensures that Minstrel never samples rates worse than the current
optimal rate unless the optimal rate experiences a failure.  See table
~\ref{table:1} for an overview of the mrr's used in Minstrel.

Minstrel's implementation is based on SampleRate's, so it also has
\texttt{ApplyRate()} and \texttt{ProcessFeedback()}, as well as a
\texttt{UpdateStats()} method that runs every
100ms.  \texttt{UpdateStats()} is home to some of the key differences
between Minstrel and SampleRate.  Instead of making decisions based on
the average transmission time, Minstrel uses
throughput: \begin{equation} T = \frac{p}{tx\_time}\end{equation}
where $T$ is the throughput of some rate $r$, $p$ is the probability
of successfully transmitting at $r$, and $tx\_time$ is the lossless
transmission time at $r$.

The probability $p$ of a successful transmission is calculated from
statistics collected by the \texttt{ProcessFeedback()} method, and is
done using an exponential weighted moving average, or EWMA.  The EWMA
creates a weighted average that weighs recent data more heavily than
old data.  The weighting for old data decreases exponentially.  An EWMA
is beneficial because it ensures that a sudden degradation in link
quality will create a rapid response in the probabilities due to the
increased weight of recent data.  Minstrel will waste less time sending
packets at a rate that no longer works, as compared to a simple
probability calculation.

\begin{table*}[htb]
  \centering
    \begin{tabular}[htb]{l|l|l|l}
    \textbf{Try} & \textbf{Normal}               & \textbf{Random (slower than best)} & \textbf{Random (faster than best)} \\ \hline
    1   & Best throughput      & Best throughput           & Random                    \\
    2   & Next best throughput & Random                    & Best throughput           \\
    3   & Highest probability  & Highest probability       & Highest probability       \\
    4   & Lowest rate          & Lowest rate               & Lowest rate               \\
    \end{tabular}
    \caption{Multi-rate retry chains for Minstrel}
\label{table:1}
\end{table*}

\section{Methodology}

We collected traces about the success rates of each bit rate through a
modified Linux wireless driver; Minstrel and SampleRate were then
re-implemented in Python and replayed on this collected data.

\subsection{Motivation}

Our initial plan was to use the algorithms as implemented in the
MadWifi drivers for the Atheros chip.  However, MadWifi is
deprecated, and only runs on now-ancient network cards.  In addition to
the unavailability of compatible hardware, MadWifi would not be an
accurate reflection of contemporary users.  Both ath5k and ath9k, which
were created to replace MadWifi, did not port the old bit-rate
selection system from MadWifi and instead opted to use the Linux
kernel's rate selection framework, which only implements Minstrel.
Porting SampleRate over to the new drivers would not have been
trivial, as there are vast differences in the assumed interfaces.

At this point, we decided to take a different approach and use an
approach similar to that used in Sprout~\cite{sprout}.  We modified
the ath9k driver to try bit rates uniformly and report packet
successes and failures, then analyze this data in user-space with
re-implementations of SampleRate and Minstrel.  Not only did this
lessen the re-implementation cost (since the algorithms were
re-implemented in user-space in Python, not in kernel-mode in C), but
it also provided reproducibility and allowed us to investigate
modifications to SampleRate and to Minstrel.

\subsection{Trace Collection}

Traces were collected from an ath9k driver modified to sample
bit-rates uniformly (forgoing the usual bit rate selection algorithm)
and to report successes and failure from each packet.  Each bit-rate
was sampled uniformly in time -- thus, more packets were sent at
higher bit-rates, since those bit-rates otherwise deliver packets
faster.  For each packet sent, the time to send was recorded, and the
bit rate for the next packet was chosen to be the bit-rate that had
the smallest total transmissions time.  This guarantees approximately
uniform selection of bit-rates.  The wireless card used was an
Qualcomm Atheros \texttt{AR9285} Wireless Network Adapter.

Traces were collected at an unused access point in a basement of MIT.
Before initiating trace collection, the wireless card was put in
monitor mode for a minute to check that no other users were connected
to the access point.  None were, guaranteeing that traces would not
include delays from carrier sensing or run into problems due to
congestion at the access point.  Traffic was simulated by sending 1500
byte packets to the access point's IP address.  Packets were sent over
UDP to avoid the TCP congestion window and retry loop; since no users
were using the access point, there was no real risk of deteriorating
service for other users.  Figure~\ref{figure:1} shows a floor plan of
the access point and basement.

Eighteen total traces were collected, grouped into a number of
scenarios.  Each trace was 30 seconds to a minute long, and each
scenario was repeated twice.  One scenario placed the wireless card
with line of sight to the access point, two meters from it.  Another
created solid body interference by having one of the authors stand
between the wireless device and the access point.  One placed the
device around a corner from the access point, creating some potential
for multi-path interference and fading.  In one scenario, we moved the
wireless device throughout the trace, pacing back and forth in a 10
meter line in front of the access point.  Finally, one five-minute
trace was taken, which mixed the above scenarios, first moving around
a corner, then standing still, then moving to line-of-sight with the
access point.  A few shorter (3 to 15 second long) traces were
collected to investigate the startup behavior of bit-rate selection.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=\linewidth]{bsmt13floorplan.png}
  \caption{We collected most traces at a remote access point in the basement of Building 13 at MIT.}
\label{figure:1}
\end{figure}

\subsection{Testing framework}

Once collected, traces were fed to a bit rate selection simulator to
analyze the performance of Minstrel and SampleRate.  Both algorithms
were represented by two functions: \texttt{ApplyRate}, which returned
a multi-rate retry chain of rates to attempt transmission at, and
\texttt{ProcessFeedback}, which was passed the success or failure of
the transmission and how many attempts were made at each bit rate
(along with the current time and the total time to send the packet);
this parallels the interface used by the Linux kernel itself.  Helper
functions, like SampleRate's \texttt{RemoveStaleResults} function,
were instead called from one of \verb|apply_rate| or
\verb|process_feedback|.  For every packet transmission, the simulator
would call \texttt{ApplyRate}; compute the probability of success
based on packets sent in 10-millisecond window around the current
time; and from that determine the success or failure of the packet,
and the number of retries required.  The results of a test run are
thus non-deterministic; both algorithms tested are non-deterministic
anyway, so this is no handicap.  To improve re-implementation
fidelity, both implementations were meticulously checked by both
authors.

We implemented SampleRate as outlined in John Bicket's Master's
thesis~\cite{samplerate}.  This implementation is slightly different
from how it was implemented in the MadWifi kernel driver.  The primary
difference is that the kernel implementatiion uses a EWMA, while the
thesis implementation computes average transmission times based on a
10-second sliding window.  It would be worth looking at the performance
of EWMA-based SampleRate in the future.  Our implementation,
\texttt{samplerate.py} contains \texttt{ApplyRate()},
\texttt{ProcessFeedback()} and \texttt{RemoveStaleResults()}, as well
as a few helper functions and data structures for tracking rate
statistics.

We implemented Minstrel by porting the C code from the 3.3.8 version
of the Linux kernel into Python.  \texttt{minstrel.py} contains
\texttt{ApplyRate()}, \texttt{ProcessFeedback()} and
\texttt{UpdateStats()}, as well as a few helper functions and data
structures for tracking rate statistics.

\subsection{Experimental parameters}

There are a few parameters that can be adjusted in Minstrel and
SampleRate, but our implementation is consistent with what the
respective authors originally chose.

In SampleRate, the number of successive retries, the frequency of
probe packets, the window size, and the lossless $tx\_time$
estimations are all parameters that can be changed.  The SampleRate
thesis chose 4 successive retries, a probe frequency of every ten
packets, and a window size of ten seconds.  We used the same values in
our implementation.  Additionally, we use the same equation to
calculate $tx\_time$.

In Minstrel, the frequency of probe packets, the window size, and the
EWMA weighting are the adjustable parameters.  The 3.3.8 Linux kernel
uses a probe frequency of 10\%, a window size of 100ms, and a EWMA
weighting of 0.75.  We use the same values.  Additionally, Minstrel also
estimates $tx\_time$ using the \texttt{ieee80211\_frame\_duration()}
method from the \texttt{util.c} file in \texttt{net/mac80211}.  We
implemented the same function in Python.

\section{Analysis} \label{sec:analysis}

Minstrel and SampleRate perform similarly, usually within 3 Mbps of
each other.  In especially noisy situations, such as when the client is
moving or when the client is around the corner from the access point,
Minstrel consistently outperforms SampleRate.  Figure~\ref{figure:3}
shows the results of a single run of both Minstrel and SampleRate over
all the different types of traces we collected.  This is consisten with
the literature---the authors of Minstrel noted that they focused on
making Minstrel robust to poor conditions~\cite{minstrel}, which means
losing potential througput gains in stable situations.  Conversely,
John Bicket noted that SampleRate performs better in stable situations
than noisy or highly varying conditions~\cite{samplerate}.  We were
slightly surprised to see how significantly SampleRate outperformed
Minstrel in certain situations, though, with throughput being up to
$67\%$ higher.

\begin{figure*}[htbp]
  \centering
  \begin{subfigure}{.45\linewidth}
    \includegraphics[width=\linewidth]{constant.png}
    \caption{Throughput of using a constant 802.11b/g bit-rate in
      clear line-of-sight from the access point.}
    \label{figure:2}
  \end{subfigure}\hfill
  \begin{subfigure}{.45\linewidth}
    \includegraphics[width=\linewidth]{MinVSSam1.png}
    \caption{Throughput of Minstrel and SampleRate in various
      scenarios.}
    \label{figure:3}
  \end{subfigure}
  \caption{Comparing Minstrel and Samplerate}
\end{figure*}

In Figure~\ref{figure:2} we show the performance of constant bit-rates
on the \texttt{clear\_1.dat} trace, which is a 35 second trace in
clear view of the access point.  The optimal rate here was 18 Mbps,
which achieved an average throughput of about 15 Mbps.  The next
highest bit-rate, 24 Mbps, shows a large plummet in performance, down
to a throughput of 1.17 Mbps.  This steep drop is charateristic, and is
also talked about in the SampleRate thesis~\cite{samplerate}.  It
demonstrates just how important bit rate selection is---the optimal
bit rate has typically much higher throughput than the next-best rate,
but if the algorithm strays too high the results can be devastating.

Bit rate selection algorithms strive to have a throughput near or
greater than the optimal constant rate.  For this
\texttt{clear\_1.dat} trace, SampleRate clearly meets this goal,
achieving an average throughput more than 1 Mbps higher than the
optimal constant rate.  In noisier situations, though, Minstrel tends
to be closer to the optimal constant rate.

Bit rate selection algorithms naturally show variation between runs on
the same data because probe rates are chosen randomly.  The variations
were small, though.  On ten runs on the \texttt{clear\_1.dat} trace,
the average of Minstrel over 10 runs was 7.52 Mbps, with a maximum
variation of +/- 0.12 Mbps.  SampleRate was similarly stable, it
averaged 5.284 Mbps over 10 runs, with a maximum variation of about
+/- 0.3 Mbps.  The variation of results on other traces (except the
very short traces, where variance was naturally higher) was similar.

The histograms of which rates the algorithms decided to send at
provided some useful insight into their behavior.  Table~\ref{table:2}
is a histogram of a 35 second corner trace.  In this situation
Minstrel outperforms SampleRate.  The constant rate throughputs
demonstrate that 11, 12 and 18 Mbps were the optimal rates to send at,
and that no packets were sucessful at any higher rates.  SampleRate
does not use these rates except for probe packets, which happens once
every ten seconds.  Minstrel, since it does not terminate transmission
after four successive failures, spends a lot of time sending probe
packets at these high rates, which are never successful.  The only
reason low rates have fewer probe packets is because Minstrel
explicitly places them second in the retry chain, and because the main
rate usually succeeds.  Despite the fact that SampleRate spends less
time probing high rates, it is less successful.  This is because
Minstrel send the most packets at the three most successful bit-rates.
SampleRate, however, spends too much time sampling low bit rates.
Additionally, it spends a lot of time sending at 5.5 Mbps.  We suspect
that this might be because the equation SampleRate uses to estimate
average transmission time needs to be tweaked.
%   TODO: why does SampleRate spend time at 5.5 Mbps???

Our examination of the histograms showed us that Minstrel spends a lot
of time sampling at high rates.  In the specific case we examined, they
were all probe packets.  But sometimes Minstrel would choose 54Mbps as
the rate with the best throughput, which made little sense.  A closer
investigation revealed that Minstrel would sometimes get lucky with a
successful probe at 54Mbps.  The probability of success would be 1, so
then 54Mbps would be chosen as the rate with the best throughput.  This
is unfortunate, but the real harm comes in the next window: packets
would now be send at 54Mbps, and nearly all would fail.  However,
since these hundreds of failed packets weight equally with the single
successful probe packet, Minstrel still considers 54 Mbps the best
rate.  Thus the probability of success might drop from 1 to about 0.70,
which is still high enough for the throughput equation to rank 54 Mbps
highly.  Fundamentally, the problem is that the EWMA does not account
the second block having hundreds more packets than the first.  In the
next section we will talk about Minproved, an improved version of
Minstrel where we attempt to fix this problem.

\section{Improvements to Minstrel}

In Section~\ref{sec:analysis}, we noted that the EWMA used by Minstrel
causes problems when probe packets succeed.  A natural next step was
to modify Minstrel to avoid this problem, by weighing each 100ms block
of statistics in proportion to how many packets it contains.  To
preserve the exponential weighing of statistics, we chose to weigh
each 100ms block based on the number of packets sent, compared to the
those sent in the average block.  Parameters were chosen so that if
all blocks were of average size, our ``balanced'' EWMA would behave
identically to a normal EWMA.

To achieve, this weighing, we first recast the usual EWMA algorithm.
The usual algorithm computes $$p \gets \alpha p + (1 - \alpha)
\frac{n_i}{d_i}),$$ where $p$ is the EWMA probability, $\alpha$ is a
weight factor equal to $0.75$ in Minstrel, and $n_i$ and $d_i$ are the
number of successful and total packets in the 100ms window.  We next
write this as $$p \gets \frac{\beta p + \frac{n_i}{d_i}}{\beta + 1},$$
where $\beta = \alpha / (1 - \alpha)$.  We can now change the factor
in front of $n_i / d_i$ to account for the total number of packets
sent.  In our case, we chose to compute $$p \gets \frac{\beta p + w_i
  \frac{n_i}{d_i}}{\beta + w_i},$$ where the weight $w_i$ is $d_i / (d
/ b)$, with $d$ the total number of packets ever send and $b$ the
total number of 100ms windows ever seen.  Then $d / b$ is the average
number of packets per window and $d_i / (d / b)$ is the ratio between
the current window and the average window.  Note that if the current
window is of average size, this is exactly equal to the usual EWMA
result.  Finally, we can simplify the above equation to $$S p \gets
\frac{\beta \frac{d}{b} p + S n_i}{\beta \frac{d}{b} + d_i},$$ where
$S$ is any constant.  This equation is usable in fixed-precision
arithmetic, making it simple to implement in the Linux kernel (which
prefers to avoid floating point computations for portability reasons).
Because packets are high rates rarely succeed, the average window size
is a few packets; thus, the main qualitative difference between our
balanced EWMA and the original Minstrel EWMA is that non-probe packets
at probe bit-rates are weighed extraordinarily heavily, and this fact
allows us to avoid multiple failing 100ms windows.

We tested a modified Minstrel which used this more balanced EWMA
algorithm on the same traces as Minstrel and SampleRate, and found a
consistent improvement of approximately 1 Mbps over the normal
Minstrel algorithm.

We noted another curious fact about Minstrel: the kernel
implementation of Minstrel samples rates very frequently.  This is due
to a flag the kernel uses to track which probe bit-rates were actually
sampled.  In cases where the probe packets are at a lower bit-rate
than the current bit-rate, the probe packets might never be sent if
the current bit-rate succeeds.  However, the flagging seems to
misbehave in certain cases, causing the extremely aggressive probing
we see in some traces.  For example, in one of the traces where the
client was around a corner from the access point, Minstrel sent almost
half of its packets at probe frequencies.  We modified Minstrel to be
less aggressive with probe packets, decreasing the sampling frequency
to approximately 10\%.  This lead to more improvement, especially on
noisy traces such as the cases where the client was around a corner or
moving.

\begin{table*}[htb]
    \centering
    \begin{tabular}{l|l|l|l|l}

    \textbf{Rate in Mbps} & \textbf{SampleRate} &
    \textbf{Minstrel} & \textbf{Minproved} &
    \textbf{Constant Rate Throughput}\\ \hline

    1   & 631  & 39    & 29    & 0.639 Mbps \\
    2   & 863  & 82    & 20    & 1.425 Mbps \\
    5.5 & 4814 & 110   & 805   & 4.406 Mbps \\
    6   & 423  & 137   & 276   & 4.603 Mbps \\
    9   & 553  & 327   & 57    & 4.630 Mbps \\
    11  & 3326 & 4121  & 3356  & 9.627 Mbps \\
    12  & 4649 & 9515  & 14285 & 9.444 Mbps \\
    18  & 1154 & 13295 & 15795 & 8.458 Mbps \\
    24  & 16   & 6909  & 1918  & 0 Mbps \\
    36  & 16   & 6944  & 2030  & 0 Mbps \\
    48  & 16   & 6888  & 2009  & 0 Mbps \\
    54  & 16   & 7308  & 1953  & 0 Mbps \\ \hline

    \textbf{Avg. Throughput:} & \emph{5.42 Mbps}  & \emph{7.578 Mbps} & \emph{11.107 Mbps} \\
    \end{tabular}
    \caption{Histograms of the number of packets sent at each rate.
      The trace was \texttt{corner\_1.dat} and was recorded around the
      corner from the access point.  The best fixed rate for this
      trace was 11 Mbps, which achieved a throughput of 9.627 Mbps.
      Minproved achieves a higher throughput than the best fixed
      rate.}

    \label{table:2}
\end{table*}

\begin{figure}[htb]
  \includegraphics[width=\linewidth]{mnVSspVSmp1.png}
  \caption{Throughput of Minproved as compared to SampleRate and
    vanilla Minstrel.}
\label{figure:4}
\end{figure}

Overall, our improved Minstrel, which we call Minproved, often
out-performs Minstrel by multiple Mbps and never doing any worse; and
usually surpassing the best constant bit rate.  In only one case did
SampleRate achieve higher throughput than Minproved, and in this case
both surpassed the best constant bit-rate.  In cases where Minstrel
out-performed SampleRate, the same behavior was true of Minproved.  In
most cases, Minproved was $30\%$ faster than Minstrel, and in some the
improvement as great as $60\%$.

However, Minproved still makes many poor choices.  While a single
successful probe packet does not cause many hundreds of milliseconds
of failed packets, as it does in Minstrel, the single successful
packet can still cause one window's worth of failed packets.  Instead,
it would be best if a successful probe packet at a rarely-successful
bit-rate was treated more carefully.  One can imagine an algorithm,
for example, which would treat a single successful probe packet as a
reason to send a dozen more probe packets at that bit-rate, but does
not yet commit to sending hundreds of packets at that bit-rate.

However, these are only vague ideas, and we have no concrete
implementations of such an algorithm.

\subsection{Availability}
All of our code and collected trace data are available on GitHub:

\noindent
{\small\url{https://github.com/pavpanchekha/6.829-project/tree/3.8.6}}

This repository contains the modified ath9k driver used to collect the
traces, the traces we collected, as well as the Python simulation
framework.

\section{Future Work}

Very little analysis has been done of the performance of bit-rate
selection algorithms on 802.11n networks.  802.11n introduces many new
rates, as well as multiple-input/multiple output, or MIMO,
capabilities.  There is an 802.11n version of Minstrel, but there is no
implementation of SampleRate.

We implemented SampleRate as outlined in the thesis, but the MadWifi
implementation has a few key differences from the thesis. MadWifi 
use of an EWMA instead of a window has more accurate
rate statistics, which will potentially lead to better throughput. We
did not have the time to implement this alternative version, but it is
worth further examination.

Hari Balakrishnan suggested that we use a broadcast approach when
collecting our traces, instead of link layer acknowledgments.  This
would entail having multiple computers listening at the broadcast
address, and compiling successes and failures from that information.
This approach would free us from certain idiosyncrasies of reading
data from the driver, and lead to possibly more accurate traces.  We
were unable to implement this due to a lack of time and ability to
acquire compatible hardware.

Finally, we are considering submitting our improvements to the
Minstrel algorithm, namely the BEWMA, as a kernel patch.  Currently is
implemented in Python, so we would have to port the code into C.

\section{Conclusion}

Minstrel and SampleRate perform similarly.  The results suggest that
SampleRate tends to do slightly better, except in very lossy links
such as the client being far awat from the AP or actively moving.
Minproved, our improved version of Minstrel, solidly outperforms both
Minstrel and SampleRate.

It seems like the abandonment of SampleRate in the Linux kernel may
have been a bit hasty, since it seems to perform similarly to Minstrel
and it is a much simpler algorithm and thus less bug prone.  We are
not sure if there are adjustments to SampleRate that will bring it's
performance near Minproved, though.

\section{Acknowledgments}

We would like to thank Jonathan Perry and Hari Balakrishnan for their
valuable guidance, as well as Derek Smithies for answering our
questions about Minstrel.

{\footnotesize \bibliographystyle{acm}
\bibliography{paper}}

\end{document}